Validating FF-Replan vs.\@ a State-Of-The-Art Probabilistic Planner in Translated Real World Domains

### 2014-03-29 FinalPaper/paper.tex
Fixing typos, adding pdf to gitignore

% first the title is needed
\title{Validating FF-Replan vs. a state-of-the-art propabilistic planner in translated real world domains}

% a short form should be given in case it is too long for the running head
\titlerunning{Presentation Plan}

@@

%
\authorrunning{Validating planners in a real world domain}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Seminar Algorithms, Embedded Software,
Msc Embedded Systems,\\
Delft University of Technology\\
\mailsa\\
}

@@

The baseline test in \cite{little2007probvsreplan} evaluates the competition problems of the IPPC described in PPDDL to correct the possible inaccurate perception of replanners vs. probabilistic planners. We would like to carefully analyze the credibility of these domains in the real world and evaluate the worth of FF-Replan vs a state of the art probabilistic planner: Gourmand \cite{kolobov2012gourmand}.

@@

\section{Domain description}
Following the theoretical comparison of probabilistic planning and replanning techniques by Little and Thiébauw, a baseline test for \emph{probabilistically interesting} domains is proposed. Problems that fail this test were deemed to be classical planning problems in disguise, favouring using replanners\cite{little2007probvsreplan}. This section describes interesting domains in which either FF-Replan or Gourmand prevailed during the International Probabilistic Planning Compeitions. We will then translate the most interesting domains to a complete "real-world" domain and evaluate them using the probabilistically interesting baseline test. This paper concludes on evaluation both planners using the translated domains.

\subsection{Interesting domains}
The competitor FF-Replan\cite{FFReplan} remembered the committee of the IPPC in 2004 that problems containing no dead ends and action costs are fairly easy to solve without much handling of uncertainty. The baseline test evaluates the competition problems of the IPPC described in PPDDLto correct the possible inaccurate perception of replanners vs. probabilistic planners.

@@

\subsection{Real world evaluation}
Explain shortcomings

\subsection{Probabilistically interesting domains}
A probabilistic planning problem is considered to be probabilistically interesting if and only if it has all the following structural properties:
\begin{itemize}
	\item there are multiple goal trajectories;
	\item there is at least on pair of distinct goal trajectories, $\tau$ and $\tau$ ', that share a common sequence of outcomes for the first n-1 outcomes, and where $\tau$\textsubscript{n} and $\tau$'\textsubscript{n} are distinct outcomes of the same action; and
	\item there exist two distinct goal trajectories $\tau$ and $\tau$' and outcome o $\in$ $\tau$ and o' $\in$ $\tau$' of two distinct actions a = act(o) and a'= act(o') such that executing a strictly decreases the maximum probability of reaching a state where a' can be executed. 
\end{itemize}

@@

\begin{itemize}
	\item Discuss the related work and how their approaches are different
	\item Discuss the other domains and why our domain adds anything to the community
	\item Introduce the current state of the art papers that use deterministic approaches for probablistic domains deadling with uncertainty
\end{itemize}

@@

\subsubsection{Reverse Iterative Deepening for Finite-Horizon MDPs with Large Branching Factors (2012)~\cite{kolobov2012glutton}}
The IPPC-2011 introduces more realistic MDPs to accomodate real-world systems with large branching factors cause by \emph{natural dynamics} and a limited time for the planner to come up with a policy. For this reason planners needed to be scalable and sensitive to the expected reward maximization. Glutton introduces an offline algorithm based on LR\textsuperscript{2}TDP and relies on the fact that a succesfull policy can be produced by solving an MDP M(H) for a smaller horizon h.  Incrementally solving the sequence M(1), M(2) forms the basis of LR\textsuperscript{2}TDP. By iterating on the distance from the horizon, data can be reused (and is stored in the cache untill the hash-table is full). The max-reward heuristic finds the largest horizon value h' for which already a heuristic value V is estimated. The  cmputation time remains much lower because it runs quickly into states that converged while solving M(h-1). The max-reward computation relies on knowing the maximum reward any action can yield in any state or an upper bound on it.

@@

\subsubsection{Probabilistic planning vs Replanning}
Inspired by the FF-Replan entrance in the IPPC competition, this paper described a probabilistic interestingness quality of domains. It then compares probabilitsic planners with replanners in the different domains to draw conclusions. Classical planning with replanning upon failure when there is so little or so much uncertainty about the world that it is not worth modelling and reasoning about or when goals dynamically change or when the problem is too large to solve at once. Probabilistic planners are more appropriate when avoiding irreparable or costly failures. Probabilistic planning computes a plan that can handle all foreseeable contingencies and is preferable for quick reactions in closed control loops.

The following structural properties are identified for comparing probabilistic planning with replanning:
\begin{itemize}
	\item 'dead' end states
	\item the degree to which the probability of reaching a deaad end state can be reduced through choice of actions
	\item number of distinct trajectories
	\item presence of mutual exclusion
\end{itemize}

If there are no dead ends, a replanner will always succeed. It is impossible to improve on the unavoidable dead ends (positive probability of reaching when executing which cannot be reduced without reucing the probability of reaching the goal).

A probabilistic planning problem is considered to be probabilistically interesting if and only if it has all the following structural properties:
\begin{itemize}
	\item there are multiple goal trajectories;
	\item there is at least on pair of distinct goal trajectories, $\tau$ and $\tau$ ', that share a common sequence of outcomes for the first n-1 outcomes, and where $\tau$\textsubscript{n} and $\tau$'\textsubscript{n} are distinct outcomes of the same action; and
	\item there exist two distinct goal trajectories $\tau$ and $\tau$' and outcome o $\in$ $\tau$ and o' $\in$ $\tau$' of two distinct actions a = act(o) and a'= act(o') such that executing a strictly decreases the maximum probability of reaching a state where a' can be executed. 
\end{itemize}

@@

They conclude by stating that the negative perception of probabilistic planning by the ICAPS community was inaccurate. They state that a synthesis of planning and replanning techniques could make a larger number of probabilistic planning problems pacticably solvable than is currently possible.

>>>>

% first the title is needed
\title{Validating FF-Replan vs.\@ a State-Of-The-Art Probabilistic Planner in Translated Real World Domains}

% a short form should be given in case it is too long for the running head
\titlerunning{FF-Replan vs Probabilistic Planner in Real World Domains}

@@

%
\authorrunning{Validating Planners in a Real World Domain}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Seminar Algorithms, Embedded Software,
MSc Embedded Systems,\\
Delft University of Technology\\
\mailsa\\
}

@@

The baseline test in \cite{little2007probvsreplan} evaluates the competition problems of the IPPC described in PPDDL to correct the possible inaccurate perception of replanners vs.\@ probabilistic planners. We would like to carefully analyze the credibility of these domains in the real world and evaluate the worth of FF-Replan vs a state of the art probabilistic planner: Gourmand \cite{kolobov2012gourmand}.

@@

\section{Domain description}
Following the theoretical comparison of probabilistic planning and replanning techniques by Little and Thi\'ebauw, a baseline test for \emph{probabilistically interesting} domains is proposed. Problems that fail this test were deemed to be classical planning problems in disguise, favouring using replanners\cite{little2007probvsreplan}. This section describes interesting domains in which either FF-Replan or Gourmand prevailed during the International Probabilistic Planning Compeitions. We will then translate the most interesting domains to a complete "real-world" domain and evaluate them using the probabilistically interesting baseline test. This paper concludes on evaluation both planners using the translated domains.

\subsection{Interesting domains}
The competitor FF-Replan\cite{FFReplan} remembered the committee of the IPPC in 2004 that problems containing no dead ends and action costs are fairly easy to solve without much handling of uncertainty. The baseline test evaluates the competition problems of the IPPC described in PPDDL to correct the possible inaccurate perception of replanners vs.\@ probabilistic planners.

@@

\subsection{Real world evaluation}

Explain shortcomings

\subsection{Probabilistically interesting domains}
A probabilistic planning problem is considered to be probabilistically interesting if and only if it has all the following structural properties:

% reference little & thiebauw again. Maybe also shortly define what all o, n, act etc. are.

\begin{itemize}
	\item there are multiple goal trajectories;
	\item there is at least on pair of distinct goal trajectories, $\tau$ and $\tau$ ', that share a common sequence of outcomes for the first n-1 outcomes, and where $\tau$\textsubscript{n} and $\tau$'\textsubscript{n} are distinct outcomes of the same action; and
	\item there exist two distinct goal trajectories $\tau$ and $\tau$' and outcome o $\in$ $\tau$ and o' $\in$ $\tau$' of two distinct actions a = act(o) and a'= act(o') such that executing a strictly decreases the maximum probability of reaching a state where a' can be executed.
\end{itemize}

@@

\begin{itemize}
	\item Discuss the related work and how their approaches are different
	\item Discuss the other domains and why our domain adds anything to the community
	\item Introduce the current state of the art papers that use deterministic approaches for probablistic domains dealing with uncertainty
\end{itemize}

@@

\subsubsection{Reverse Iterative Deepening for Finite-Horizon MDPs with Large Branching Factors (2012)~\cite{kolobov2012glutton}}
The IPPC-2011 introduces more realistic MDPs to accommodate real-world systems with large branching factors cause by \emph{natural dynamics} and a limited time for the planner to come up with a policy. For this reason planners needed to be scalable and sensitive to the expected reward maximization. Glutton introduces an off-line algorithm based on LR\textsuperscript{2}TDP and relies on the fact that a successful policy can be produced by solving an MDP M(H) for a smaller horizon h.  Incrementally solving the sequence M(1), M(2) forms the basis of LR\textsuperscript{2}TDP. By iterating on the distance from the horizon, data can be reused (and is stored in the cache until the hash-table is full). The max-reward heuristic finds the largest horizon value h' for which already a heuristic value V is estimated. The  computation time remains much lower because it runs quickly into states that converged while solving M(h-1). The max-reward computation relies on knowing the maximum reward any action can yield in any state or an upper bound on it.

@@

\subsubsection{Probabilistic planning vs Replanning}
Inspired by the FF-Replan entrance in the IPPC competition, this paper described a probabilistic interestingness quality of domains. It then compares probabilistic planners with replanners in the different domains to draw conclusions. Classical planning with replanning upon failure when there is so little or so much uncertainty about the world that it is not worth modelling and reasoning about or when goals dynamically change or when the problem is too large to solve at once. Probabilistic planners are more appropriate when avoiding irreparable or costly failures. Probabilistic planning computes a plan that can handle all foreseeable contingencies and is preferable for quick reactions in closed control loops.

The following structural properties are identified for comparing probabilistic planning with replanning:
\begin{itemize}
	\item 'dead' end states
	\item the degree to which the probability of reaching a dead end state can be reduced through choice of actions
	\item number of distinct trajectories
	\item presence of mutual exclusion
\end{itemize}

If there are no dead ends, a replanner will always succeed. It is impossible to improve on the unavoidable dead ends (positive probability of reaching when executing which cannot be reduced without reducing the probability of reaching the goal).

% repetition of what's already stated earlier

A probabilistic planning problem is considered to be probabilistically interesting if and only if it has all the following structural properties:

\begin{itemize}
	\item there are multiple goal trajectories;
	\item there is at least on pair of distinct goal trajectories, $\tau$ and $\tau$ ', that share a common sequence of outcomes for the first n-1 outcomes, and where $\tau$\textsubscript{n} and $\tau$'\textsubscript{n} are distinct outcomes of the same action; and
	\item there exist two distinct goal trajectories $\tau$ and $\tau$' and outcome o $\in$ $\tau$ and o' $\in$ $\tau$' of two distinct actions a = act(o) and a'= act(o') such that executing a strictly decreases the maximum probability of reaching a state where a' can be executed.
\end{itemize}

@@

They conclude by stating that the negative perception of probabilistic planning by the ICAPS community was inaccurate. They state that a synthesis of planning and replanning techniques could make a larger number of probabilistic planning problems practicably solvable than is currently possible.
